---
title: " Metadata exploration"
author: "Florentin Constancias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
rm(list = ls())
gc()
# options(java.parameters = "-Xmx80000m")
```

```{r setup2, include=FALSE}
##knitr::opts_chunk$set(cache=TRUE)
##knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(dev = "ragg_png") # https://ragg.r-lib.org/index.html
##knitr::opts_chunk$set(fig.width = 6) #https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing
knitr::opts_chunk$set(fig.asp = 0.618)
knitr::opts_chunk$set(fig.show = "hold")
knitr::opts_chunk$set(fig.show = "70%")
knitr::opts_chunk$set(fig.align = "center")
```

```{r setup3, include=FALSE}
library(tidyverse)
library(phyloseq)

library(readxl)
require(microbiomeMarker)
```


```{r}
load(here::here("phyloseq.RData"))
```




slow moderqte rqpid
```{r}
sample_data(physeq) %>% data.frame() %>% rownames_to_column("name") -> raw_fasta_names

raw_fasta_names

raw_fasta_names %>% 
  write_tsv("niko_samples.tsv")
```

```{r}
# source("https://raw.githubusercontent.com/fconstancias/metabaRpipe-source/master/Rscripts/functions.R")
# 
# physeq %>%
#   physeq_add_metadata(physeq = .,
#                       metadata = "../../data-raw/ps_invivo_meta.xlsx" %>%
#                         readxl::read_xlsx(),
#                       sample_column = "Sample") 
```

```{r}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R") 

physeq %>% 
  generate_color_palette(var = "Group",
                         pal = "npg",
                         print = FALSE) -> treat_pal


physeq %>% 
  generate_color_palette(var = "time",
                         pal = "jco",
                         print = FALSE) -> time_pal
```

# Sample to download from ENA:

```{r}
# "/Users/florentin/Documents/Nikoline/01_RawData/ena-file-download-read_run-PRJEB69273-submitted_ftp-20240424-1900.sh" %>% 
#   read.table() %>% 
#   # mutate("V4" = str_extract(V3, "\\w+$"))
#   # mutate("V4" = sub(".*/", "", V3))
#   mutate("V4" = basename(df$V3)) %>% 
#   mutate("V5" =    str_extract(V4,pattern = "[^_]+")) %>% 
#   # mutate("V6" =    str_extract(V4,pattern = "[^-]+")) %>% 
#   mutate("V6" = gsub("-", "", V5)) -> ena
```

```{r}
# ena %>% 
#   select(V3, V4,V6) %>% 
#   right_join(raw_fasta_names %>% 
#                select(name, time), 
#              join_by("V6" == "name"),
#              keep = TRUE) %>% 
#   arrange(V3) -> dl_ready
# 
# dl_ready
# 
# 
# dl_ready %>% 
#   write_tsv("/Users/florentin/Documents/Nikoline/01_RawData/dlreqdy.tsv")
```


# Metadata:

# Visualisation and beta-dive evaluation per time point responder groups:

# Rarefaction curves/

```{r}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 

physeq %>%  sample_sums() %>%  min() -> min_lib

min_lib
```


```{r}
physeq %>%
  rarefy_even_depth(rngseed = 123,
                    sample.size = min_lib) -> physeq_rare # important to specify rngseed: random number generator for reproducibility
```
```{r}

physeq_rare %>%
  phyloseq_rarefaction_curves(stepsize = 100, 
                              color_data = "Group", 
                              facet_data = NULL) -> pr

pr + geom_vline(xintercept = min_lib,
                color = "red",
                linetype = "dashed", size=0.25) +
  facet_wrap(. ~ time) + ylab("ASV Richness") -> plot_rare

plot_rare
```

# Beta-diversity:

## compute distances:

```{r}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R") 

physeq_rare %>% 
  microbiome::transform(., "hellinger") %>% 
  phyloseq::distance(method = "bray") -> bc
```

```{r}
physeq_rare %>% 
  phyloseq_compute_bdiv(phylo = FALSE) -> dlist
```

```{r}
physeq %>% 
  microViz::dist_calc(., dist = "aitchison") %>% 
  microViz::dist_get() %>% magrittr::divide_by(100) -> aitch
```

```{r}
dlist$bray <- bc
dlist$aitch <- aitch
```

```{r}
plot(dlist$bray, dlist$aitch)
```

```{r}
cor.test(dlist$bray, dlist$aitch)
```
## Ordination:

```{r}
physeq_rare %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```
```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```
```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

### Selected ordination with trqjectories and envfit taxa:

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = Group ,
                                 shape = time,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```

## Statistical evalutaion:

### Overall:

```{r}
form = paste0(c("Group", "time"), collapse=" * ")

lapply(
  dlist,
  FUN = phyloseq_adonis2,
  physeq = physeq_rare,
  formula = form,
  nrep = 999,
  # strata = "id"
)  %>%
  bind_rows(.id = "Distance") %>% 
  mutate("Group" = (as.vector(form))) -> perm1

perm1 %>% 
  mutate_if(is.numeric, round, 3) %>%
  filter(! terms %in% (c("Residual", "Total"))) %>%
  DT::datatable()
```

```{r}
form = paste0(c("Group"), collapse=" * ")


lapply(
  dlist,
  FUN = phyloseq_adonis2,
  physeq = physeq_rare,
  formula = form,
  nrep = 999,
  # strata = "id"
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate("Group" = (as.vector(form))) -> perm2

perm2 %>% 
  mutate_if(is.numeric, round, 3) %>%
  filter(! terms %in% (c("Residual", "Total"))) %>%
  DT::datatable()
```

```{r}
form = paste0(c("time"), collapse=" * ")

lapply(
  dlist,
  FUN = phyloseq_adonis2,
  physeq = physeq_rare,
  formula = form,
  nrep = 999,
  # strata = "id"
)  %>%
  bind_rows(.id = "Distance") %>% 
  mutate("Group" = (as.vector(form))) -> perm3

perm3 %>% 
  mutate_if(is.numeric, round, 3) %>%
  filter(! terms %in% (c("Residual", "Total"))) %>%
  DT::datatable()
```

```{r}

form = paste0(c("Group", "time"), collapse=" + ")

lapply(
  dlist,
  FUN = phyloseq_adonis2,
  physeq = physeq_rare,
  formula = form,
  nrep = 999,
  # strata = "id"
)    %>%
  bind_rows(.id = "Distance") %>% 
  mutate("Group" = (as.vector(form))) -> perm4

perm4 %>% 
  mutate_if(is.numeric, round, 3) %>%
  filter(! terms %in% (c("Residual", "Total"))) %>%
  DT::datatable()
```


```{r}
bind_rows(perm1, perm2,perm3, perm4) -> combined_perm

combined_perm
```


### Pairwise:

#### Within time points:

##### T1:

```{r}
physeq_rare %>% 
  subset_samples(time == "T1") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "Group",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate(group = "T1") -> pwperm_t1

pwperm_t1 %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```

```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = Group ,
                                 shape = time,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```


```{r}
physeq_tmp %>% 
  run_lefse(., group = "Group", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2, multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```

##### T2:

```{r}
physeq_rare %>% 
  subset_samples(time == "T2") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "Group",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>%
  mutate(group = "T2") -> pwperm_t2

pwperm_t2 %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```


```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = Group ,
                                 shape = time,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```


```{r}
physeq_tmp %>% 
  run_lefse(., group = "Group", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2 ,multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```



##### T3:

```{r}
physeq_rare %>% 
  subset_samples(time == "T3") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "Group",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate(group = "T3") -> pwperm_t3

pwperm_t3 %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```

```{r}
physeq_tmp %>% 
  run_lefse(., group = "Group", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2, multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```


```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = Group ,
                                 shape = time,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```

#### Within groups:

##### Slow:


```{r}
physeq_rare %>% 
  subset_samples(Group  == "1. Slow") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "time",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate(group = "Intermediate") -> pwperm_slow

pwperm_slow %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```

```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = time ,
                                 shape = NULL,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```

```{r}
physeq_tmp %>% 
  run_lefse(., group = "time", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2, multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```



##### Intermediate:

```{r}
physeq_rare %>% 
  subset_samples(Group == "2. Intermediate") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "time",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate(group = "Intermediate") -> pwperm_int

pwperm_int %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```

```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = time ,
                                 shape = NULL,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```

```{r}
physeq_tmp %>% 
  run_lefse(., group = "time", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2, multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```


##### Fast:

```{r}
physeq_rare %>% 
  subset_samples(Group == "3. Fast") -> physeq_tmp

lapply(
  dlist,
  FUN = physeq_pairwise_permanovas_adonis2,
  physeq = physeq_tmp,
  compare_header = "time",
  n_perm = 999,
  strat = FALSE
) %>%
  bind_rows(.id = "Distance") %>% 
  mutate(group = "Fast") -> pwperm_fst

pwperm_fst %>% 
  mutate_if(is.numeric, round, 3) %>%
  # filter(! terms %in% (c("Residuals", "Total"))) %>%
  DT::datatable()
```


```{r}
physeq_tmp %>% 
  phyloseq_plot_bdiv(dlist = dlist, # list of distance computed from a phyloseq object
                     ps_rare = ., # phyloseq object
                     m = "PCoA", # PCoA or NMDS
                     seed = 123, # for reproducibility
                     axis1 = 1, # axis to plot
                     axis2 = 2) -> plot_list
```

```{r}
plot_list %>%
  phyloseq_plot_ordinations_facet(color_group = "Group",
                                  shape_group = "time",
                                  alpha = NULL) -> pcoas

pcoas
```

```{r}
plot_list %>%
  phyloseq_ordinations_expl_var() %>%
  DT::datatable()
```

```{r}
plot_list$wjaccard$layers[[1]] = NULL; plot_list$wjaccard$layers[[1]] = NULL
plot_list$wjaccard$layers[[2]] = NULL; plot_list$wjaccard$layers[[1]] = NULL

# plots_hall_humans$aichinson$layers[[1]] = NULL;plots_hall_humans$aichinson$layers[[1]] = NULL
# plots_hall_humans$aichinson$layers[[2]] = NULL;plots_hall_humans$aichinson$layers[[2]] = NULL

plot_list$wjaccard + geom_point(size = 3,
                             aes(colour = time ,
                                 shape = NULL,
                                 alpha = NULL)) +
  geom_path(data = plot_list$wjaccard$data %>%
              arrange(Group) ,
            # aes(colour = Treatment, group = interaction(Model, Model2, Antibiotic, Treatment, Fermentation, Reactor,Antibiotic_mg.L)),
            aes(colour = Group, group = id),
            
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black") +
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "right")  -> PCOA

PCOA
```

```{r}
PCOA + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 3))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp

physeq_rare %>%
  phyloseq_add_taxa_vector_fix(phyloseq = .,
                               dist = dlist$wjaccard,
                               taxrank_glom = "Family",
                               figure_ord = empty_plot_tmp, 
                               fact = 0.2, pval_cutoff = 0.05,
                               top_r = 8) -> out #top 10 correlated genera

out$signenvfit %>% 
  DT::datatable()
```

```{r}
out$vectors
```

```{r}
physeq_tmp %>% 
  run_lefse(., group = "time", norm = "CPM",
            kw_cutoff = 0.05, lda_cutoff = 2, multigrp_strat = FALSE, strict = "0") -> lef_out
```

```{r}
# physeq_tmp %>% 
#   run_ancombc(., group = "Group", norm = "CPM") -> ancombc_out
# 
# ancombc_out

# Warning: No marker was identified
```

```{r}
# physeq_tmp %>%
#   run_ancom(., group = "Group", norm = "CPM") -> ancom_out
# 
# ancom_out
# 
# no microbiome markers were identified 
```

```{r}
lef_out %>% 
  marker_table() %>%  data.frame() -> lefse_df

```

```{r}
plot_ef_bar(lef_out) +
  scale_color_manual(name = "", values = time_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = time_pal,
                    na.value = "black")  -> lef_plot

lef_plot
```

Combine results:

```{r}
bind_rows(pwperm_fst, pwperm_int, pwperm_slow, pwperm_t1, pwperm_t2, pwperm_t3) -> combined_pw_perm

combined_pw_perm
```

### Plot Distqnce to bqseline from qll the responders


```{r}
physeq_rare %>% 
  microViz::ps_mutate(tmp2 = paste0(id,"_", time,"_", Group, "_", "wjaccard")) %>% sample_data() %>% 
  phyloseq_distance_boxplot(p = ., # phyloseq object
                            dist = dlist$wjaccard, # distance matrix
                            d = "tmp2") -> outwjaccard # column in metadata for comparaisons

physeq_rare %>% 
  microViz::ps_mutate(tmp2 = paste0(id,"_", time,"_", Group, "_", "aitch")) %>% sample_data() %>% 
  phyloseq_distance_boxplot(p = ., # phyloseq object
                            dist = dlist$aitch, # distance matrix
                            d = "tmp2") -> outaitch # column in metadata for comparaisons



outwjaccard$matrix %>% 
  mutate(distance = "wjaccard") %>% 
rbind(., 
   outaitch$matrix %>% 
  mutate(distance = "aitch")) %>% 
  separate(Type1, into = c("id_1", "time_1", "group_1", "dist_1") , sep = "_") %>% 
  separate(Type2, into = c("id_2", "time_2", "group_2", "dist_2") , sep = "_") -> combined


combined %>% 
  filter(id_1 == id_2,
         time_1 == "T1",
         time_2 != "T1",
         dist_1 == dist_2) %>% 
  ggplot(data = ., aes(x=group_2, y=value)) +
  geom_boxplot(outlier.colour = NA, alpha=0.7, aes(fill = group_2, color = group_2)) +
  # ggbeeswarm::geom_beeswarm(size=1, alpha=0.2,
  #                           position=position_jitterdodge(dodge.width=0.9)) +
  geom_jitter(size=1, position = position_jitterdodge(dodge.width=1),
              aes_string(color = "group_2")) + 
  facet_grid(as.formula(paste0("dist_1 ~ ",paste("time_2"))), scales = "free_y", space = "fixed") +
  scale_color_manual(name = "", values = treat_pal,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_pal,
                    na.value = "black") -> p 
  # scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +

p
```


```{r}
  ggpubr::compare_means(formula = as.formula(paste0("value ~ ", paste0("time_2"))),
                        group.by = c("dist_1", "group_1"),
                        data = p$data,
                        method = "wilcox.test",
                        p.adjust.method = "fdr") %>%
    select(-.y., -p.format, -p.signif) %>%
    arrange(p) %>%
    mutate(signif = ifelse(p.adj <= 0.05, 'SIGN', 'NS')) -> dist_box_stats

dist_box_stats %>% 
  DT::datatable()
```


```{r}
save(physeq, treat_pal, time_pal, physeq_rare ,alpha_save, plot_rare, heat_all,file = here::here("save.RData"))

```

```{r}
sessionInfo()
```

